---
title: "CASA0005_week4_homework"
author: "Chao Sun"
output: md_document
date: "2024-10-25"
---

# week4_homework

## Load necessary packages

Loads necessary libraries for data reading, spatial analysis, data manipulation, and visualization.

```{r setup, include=FALSE}
library(readr)
library(sf)
library(here)
library(countrycode)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Reading Data

Loads datasets containing composite indices and world country boundaries, then retrieves their column names.

```{r}
# Load data
HDR_Composite_Data <- read.csv(here("week4_homework_data", "HDR23-24_Composite_indices_complete_time_series.csv"))
World_Map <- st_read(here("week4_homework_data", "World_Countries_(Generalized).geojson"))

# Check column names
colnames(HDR_Composite_Data)
colnames(World_Map)
```

## Prepare and merge data

Calculates the change in Gender Inequality Index (GII) between 2010 and 2019 and merges it with a world countries dataset by converting country codes.

```{r}
# Extract Gender Inequality Index (GII) data and calculate the difference between 2010 and 2019
gender_inequality_index <- HDR_Composite_Data %>%
  select(iso3, contains("gii")) %>%
  mutate(gii_change_2019_2010 = gii_2019 - gii_2010)

# Process geographic data and merge it with gender inequality data
World_Map_Modified <- World_Map %>%
  mutate(ISO3_Code = countrycode(ISO, "iso2c", "iso3c")) %>%
  left_join(gender_inequality_index, by = c("ISO3_Code" = "iso3"))
```

## Plot a map

Visualizes the difference in Gender Inequality Index (GII) between 2010 and 2019 across countries.

```{r}
# Visualize the difference in the Gender Inequality Index
ggplot(World_Map_Modified) + 
  geom_sf(aes(fill = gii_change_2019_2010), 
          color = "#A9A9A9",  
          size = 0.1) + 
  scale_fill_gradient2(
    low = "#FDB927", 
    mid = "#FFFFFF", 
    high = "#552583", 
    midpoint = 0, 
    na.value = "#D3D3D3", 
    name = "GII Difference") + 
  theme_minimal() + 
  labs(title = "Global Gender Inequality Index Difference 2010-2019") + 
  theme(plot.title = element_text(hjust = 0.5))
```
